Finiteness in Cubical Type Theory

### 2020-01-21 paper.tex
paper
definitions done

All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in the
final section, where we implement a library for proof search, based on
omniscience and exhaustibility.
While proof search based on finiteness is not new, implementing it here does
give us an opportunity to demonstrate some actual \emph{computation} in a
univalent setting.
Many of the finiteness predicates are built using univalence, so for the proof
search functionality to work at all we require computational content in the
univalence axiom.
Furthermore, the extensionality afforded to us by HoTT means that we can have
\(\Pi\) types in the domain of the search, meaning we can synthesise full
functions from specifications alone.
\subsection{Related Work}
\subsection{Notation and Background}
\subsubsection{Notation}
We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
For the various type formers we use the following notation:
\begin{description}
  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
    ``Type families'' are functions into type.
  \item[Universes]
    Implicitly, types will exist at some point in a universe hierarchy,
    beginning with \(\mathbf{Type}\) (\(\mathbf{Type} : \mathbf{Type}_1\),
    \(\mathbf{Type}_1 : \mathbf{Type}_2\), and so on).
    In our formalisation, every proof is done in the most universe polymorphic
    way possible: the proof that the cardinally finite sets form a
    \(\Pi\)-pretopos, for instance, is defined over any universe level.
  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
    The single inhabitant of \(\top\) is tt, and the two inhabitants of
    \(\mathbf{Bool}\) are false and true.
  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
    dependent sum and product, respectively.
    The two projections from \(\Sigma\) are called fst and snd.
    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
    \(\Pi\) as \(\rightarrow\).
  \item[Disjoint Union] Disjoint union can be defined in terms of \(\Sigma\):
    \begin{equation}
      A \uplus B \coloneqq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then } A \text{ else } B
    \end{equation}
    However, we prefer to use it as an inductively defined type (though the two
    are equivalent).
    The two constructors are:
    \begin{equation}
      \begin{aligned}
        \text{inl} &: A \rightarrow A \uplus B \\
        \text{inr} &: B \rightarrow A \uplus B
      \end{aligned}
    \end{equation}
  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
    for definitions.
    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
    so the distinction isn't terribly important in usage.
\end{description}
\subsubsection{Cubical Type Theory}
Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
theory with an implementation in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
It allows us to do much of the same theory as in HoTT, but crucially the
univalence ``axiom'' is a \emph{theorem}, rather than an axiom.
This allows us to actually compute with univalent proofs, a capability missing
from HoTT.
\begin{romdefinition}[Path Types]
  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
  Paths\footnotemark.
  The internal structure of paths is largely irrelevant to us here, as we will
  generally treat \(\equiv\) as a black-box equivalence relation with
  substitution and congruence.
\end{romdefinition}

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{romdefinition}[Homotopy Levels]
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  This means that equalities are not necessarily unique.
  In fact, we have an entire hierarchy of homotopy levels, of which sets are
  level 2.
  \begin{align}
    \text{isContr}(A)    &\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
    \text{isProp}(A)     &\coloneqq \Pi(x, y : A) , (x \equiv y) \\
    \text{isSet}(A)      &\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
    \text{isGroupoid}(A) &\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
  \end{align}
  We can define the above types inductively like so:
  \begin{align}
    \text{isOfHLevel}(0 , A) &\coloneqq \text{isContr}(A) \\
    \text{isOfHLevel}(n+1 , A) &\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
  \end{align}

@@

\begin{romdefinition}[Isomorphism]
  We say two types \(A\) and \(B\) are isomorphic, denoted with \(A \iff B\), if
  there is a pair of functions \(f : A \rightarrow B\), \(g : B \rightarrow A\)
  such that
  \begin{equation}
    \begin{aligned}
      \Pi(x : A) , (g(f(x)) \equiv x) \\
      \Pi(x : B) , (f(g(x)) \equiv x)
    \end{aligned}
  \end{equation}
  While we can derive an equivalence from an isomorphism:
  \begin{equation}
    \text{isoToEquiv} : (A \iff B) \rightarrow (A \simeq B)
  \end{equation}
  It is not the case that an isomorphism is the \emph{same} as an equivalence in
  all cases.
  In other words, we do not have:
  \begin{equation}
    (A \iff B) \simeq (A \simeq B)
  \end{equation}
  However, if one of \(A\) or \(B\) is a set, then the above holds.
\end{romdefinition}

% \item[Isomorphisms and Biconditionals] A pair of functions \(f : A \rightarrow
%   B\) and \(g : B \rightarrow A\) will be described by the type \(A
%   \leftrightarrow B\).
%   If \(g\) is a quasi-inverse \cite[definition 2.4.6]{hottbook} of \(f\), we
%   have an isomorphism between \(A\) and \(B\), denoted \(A \iff B\).
%   Note that while we can derive an equivalence from an isomorphism, we do not
%   have \((A \iff B) \simeq (A \simeq B)\).

We will use the standard \(\Pi\) and \(\Sigma\) notation for dependent products
and sums, respectively.

\todo[inline]{HITs}
\todo[inline]{Surjections}
\todo[inline]{Containers}
\todo[inline]{Fin}
\todo[inline]{Biconditional, Isomorphism, Equivalence, Equals}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Split Enumerable}
\begin{romdefinition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\todo[inline]{Put refs for each ``definition of''}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    & \text{Definition of } \mathcal{E}!
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    & \text{Definition of } \in
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of sp-surj}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}

@@

Split enumerability implies decidable equality on the underlying type.
To prove this, we will make use of the following lemma, proven in the
formalisation:
\begin{romlemma} \label{discrete-surj}
  \begin{equation}
    \frac{
        A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
      }{
       \text{Discrete}(B) 
      }
  \end{equation}
\end{romlemma}
\begin{romlemma}
  Every split enumerable type is discrete.
\end{romlemma}
\begin{proof}
  Let \(A\) be a split enumerable type.
  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
  \(\mathbf{Fin}(n)\) for some \(n\).
  Also, by lemma \todo[noline]{Fin is discrete}, we know that
  \(\mathbf{Fin}(n)\) is discrete.
  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
  \qed
\end{proof}
\subsection{Manifest Bishop}
\begin{romdefinition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
  \end{equation}
\end{romdefinition}
The only difference between manifest Bishop finiteness and split enumerability
is the membership term: here we require unique membership (\(\inunique\)),
rather than simple membership (\(\in\)).

@@

% Before giving the definition of the predicate, we will first define lists (and
% membership thereof): we have chosen a \emph{container}-based definition for this
% work.
% \begin{romdefinition}[Container] \label{container-def}
%   A container \cite{abbottContainersConstructingStrictly2005} is a pair
%   \(S \triangleright P\) where \(S\) is a type, the elements of which are called
%   the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
%   the elements of \(P(s)\) are called the \emph{positions} of a container.
%   We ``interpret'' a container into a functor defined like so:
%   \begin{equation}
%     \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
%   \end{equation}
%   Membership of a container can be defined like so:
%   \begin{equation}
%     x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
%   \end{equation}
%   Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
% \end{romdefinition}
% \begin{romdefinition}[Lists]
%   \begin{equation}%
%     \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
%   \end{equation}
% \end{romdefinition}
% \begin{romdefinition}[\mathbf{Fin}]
%   \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
%   We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
%   \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
% \end{romdefinition}

@@

% Finally, we can define formally split enumerability.
% \subsection{Split Surjections}
% \begin{romdefinition}[Surjections]
%   \begin{multicols}{2}
%     \begin{equation}
%       A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
%     \end{equation} \break
%     \begin{equation}
%       A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
%     \end{equation}
%   \end{multicols}
%   Where the definitions of surjective are taken from \cite[definition
%   4.6.1]{hottbook}.
% \end{romdefinition}
% \begin{romtheorem} \label{split-enum-surj}
%   Split enumerability is equivalent to a split surjection from a finite prefix
%   of the natural numbers.
%   \begin{equation}
%     \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   After sufficient inlining, it emerges that our goal is simply a reassociation.
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
%   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
%   have written out the extra lines for clarity alone.
% \end{proof}
% \begin{romlemma} \label{split-enum-discrete}
%   Any split enumerable type has decidable equality (is discrete).
% \end{romlemma}
% \begin{proof}
%   We use a corollary that if there is a split-surjection from \(A\) to \(B\),
%   and \(A\) is discrete, then \(B\) is also discrete.
% \end{proof}
% By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
% this implies split enumerable types must all be sets.
% \subsection{Closure}
% In this section we will prove closure under various operations for split
% enumerable sets.
% We are working towards a topos proof, which requires us to prove closure under
% a variety of operations: for now, we only have enough machinery to demonstrate
% the semiring operations, and dependent sums.
% In order to show closure under exponentials (function arrows), we will need an
% equivalence with \(\mathbf{Fin}\), which will be provided in
% section~\ref{manifest-bishop}.
% \begin{romlemma} \label{split-enum-bool-top-bot}
%   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
% \end{romlemma}
% \begin{proof}
%   Each of these types clearly has a finite number of elements (2, 1, and 0,
%   respectively), and furthermore has a straightforward enumeration.
% \end{proof}
% \begin{romtheorem} \label{split-enum-sigma}
%   Split-enumerability is closed under \(\Sigma\).
%   \begin{equation} \label{split-enum-sigma-clos-sig}
%     \frac{
%       \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
%     }{
%       \mathcal{E}!(\Sigma (x : A) , U(x))
%     }
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   To obtain the support list, we concatenate the support lists of all the proofs
%   of split-finiteness for \(U\) over the support list of \(E_A\).
%   In Agda:
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
%   ``do-notation'' is available to us as we're working in the list monad.
% \end{proof}
% Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
% as these types can be defined in terms of \(\Sigma\).
% \begin{align}
%   A \times B &= \Sigma {(x : A)} , B \\
%   A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
% \end{align}
% \section{Manifest Bishop Finiteness} \label{manifest-bishop}
% Where split enumerability was the enumeration form of a split surjection from
% \(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
% \emph{equivalence} with \(\mathbf{Fin}\).

>>>>

All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in the
final section, where we implement a library for proof search, based on
omniscience and exhaustibility.
While proof search based on finiteness is not new, implementing it here does
give us an opportunity to demonstrate some actual \emph{computation} in a
univalent setting.
Many of the finiteness predicates are built using univalence, so for the proof
search functionality to work at all we require computational content in the
univalence axiom.
Furthermore, the extensionality afforded to us by HoTT means that we can have
\(\Pi\) types in the domain of the search, meaning we can synthesise full
functions from specifications alone.
\subsection{Related Work}
\subsection{Notation and Background}
\subsubsection{Notation}
We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
For the various type formers we use the following notation:
\begin{description}
  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
    ``Type families'' are functions into type.
  \item[Universes]
    Implicitly, types will exist at some point in a universe hierarchy,
    beginning with \(\mathbf{Type}\) (\(\mathbf{Type} : \mathbf{Type}_1\),
    \(\mathbf{Type}_1 : \mathbf{Type}_2\), and so on).
    In our formalisation, every proof is done in the most universe polymorphic
    way possible: the proof that the cardinally finite sets form a
    \(\Pi\)-pretopos, for instance, is defined over any universe level.
  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
    The single inhabitant of \(\top\) is tt, and the two inhabitants of
    \(\mathbf{Bool}\) are false and true.
  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
    dependent sum and product, respectively.
    The two projections from \(\Sigma\) are called fst and snd.
    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
    \(\Pi\) as \(\rightarrow\).
  \item[Disjoint Union] Disjoint union can be defined in terms of \(\Sigma\):
    \begin{equation}
      A \uplus B \coloneqq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then } A \text{ else } B
    \end{equation}
    However, we prefer to use it as an inductively defined type (though the two
    are equivalent).
    \begin{equation}
      \begin{aligned}
        A \uplus B \coloneqq & \;
        \text{inl} &: A \rightarrow A \uplus B \\
        | & \;  \text{inr} &: B \rightarrow A \uplus B
      \end{aligned}
    \end{equation}
  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
    for definitions.
    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
    so the distinction isn't terribly important in usage.
\end{description}
\subsubsection{Cubical Type Theory}
Cubical Type Theory \cite{cohenCubicalTypeTheory2016} is a constructive type
theory with an implementation in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
It allows us to do much of the same theory as in HoTT, but crucially the
univalence ``axiom'' is a \emph{theorem}, rather than an axiom.
This allows us to actually compute with univalent proofs, a capability missing
from HoTT.
\begin{romdefinition}[Path Types] \label{path-types}
  The equality type (which we denote with \(\equiv\)) in CuTT is the type of
  Paths\footnotemark.
  The internal structure of paths is largely irrelevant to us here, as we will
  generally treat \(\equiv\) as a black-box equivalence relation with
  substitution and congruence.
\end{romdefinition}

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{romdefinition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  This means that equalities are not necessarily unique.
  In fact, we have an entire hierarchy of homotopy levels, of which sets are
  level 2.
  \begin{align}
    \text{isContr}(A)    &\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
    \text{isProp}(A)     &\coloneqq \Pi(x, y : A) , (x \equiv y) \\
    \text{isSet}(A)      &\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
    \text{isGroupoid}(A) &\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
  \end{align}
  We can define the above types inductively like so:
  \begin{align}
    \text{isOfHLevel}(0 , A) &\coloneqq \text{isContr}(A) \\
    \text{isOfHLevel}(n+1 , A) &\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
  \end{align}

@@

\begin{romdefinition}[Fibers] \label{fibers}
  A fiber is defined over some function \(f : A \rightarrow B\).
  \begin{equation}
    \text{fib}_f(y) = \Sigma(x : A) , (f x \equiv y)
  \end{equation}
\begin{romdefinition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{equation}
    A \simeq B \coloneqq \Sigma(f : A \rightarrow B) , \Pi(y : B) , \text{isContr}(\text{fib}_f(y))
  \end{equation}
\begin{romlemma}
  Univalence
  \begin{equation}
    (A \simeq B) \simeq (A \equiv B)
  \end{equation}
\end{romlemma}
\begin{romdefinition}[Isomorphism] \label{isomorphism}
  We say two types \(A\) and \(B\) are isomorphic, denoted with \(A \iff B\), if
  there is a pair of functions \(f : A \rightarrow B\), \(g : B \rightarrow A\)
  such that
  \begin{equation}
    \begin{aligned}
      \Pi(x : A) , (g(f(x)) \equiv x) \\
      \Pi(x : B) , (f(g(x)) \equiv x)
    \end{aligned}
  \end{equation}
  While we can derive an equivalence from an isomorphism:
  \begin{equation}
    \text{isoToEquiv} : (A \iff B) \rightarrow (A \simeq B)
  \end{equation}
  It is not the case that an isomorphism is the \emph{same} as an equivalence in
  all cases.
  In other words, we do not have:
  \begin{equation}
    (A \iff B) \simeq (A \simeq B)
  \end{equation}
  However, if one of \(A\) or \(B\) is a set, then the above holds.
\begin{romdefinition}[Higher Inductive Types] \label{HITs}
  \todo[inline]{define}
\begin{romdefinition}[Surjections] \label{surjections}
  \begin{alignat}{3}
    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \\
    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \\
    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
  \end{alignat}
\begin{romdefinition}[Containers] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S , P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{equation} \label{container-interp}
    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation} \label{container-membership}
    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
  \end{equation}
\begin{romdefinition}[\(\mathbf{List}\)] \label{List}
  \begin{equation}
    \mathbf{List} \coloneqq \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
  \end{equation}
\begin{romdefinition}[\(\mathbf{Fin}\)] \label{Fin}
  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
  We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
  \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
\begin{romdefinition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

  It is defined as a Higher Inductive Type:
  \begin{equation} {
    \begin{alignat*}{3}
      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &&: A \rightarrow \lVert A \rVert ; \\
                              | & \; \text{squash}     &&: \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{alignat*} }
  \end{equation}
  We will use three eliminators from \(\lVert A \rVert\) in this paper.
  \begin{enumerate}
  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
  \item \label{elim-prop-monad} A special case of \ref{elim-prop-prop} implies
    that \(\lVert \cdot \rVert\) forms a monad: this means that we get the usual
    Monadic operators on \(\lVert \cdot \rVert\) (bind, pure, fmap, etc.).
    \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
      function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
      choice of \(A\) (\(\Pi {(x , y : A)} , f(x) \equiv f(y) \)).
      Formally speaking, \(f\) needs to be ``coherently constant''
      \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
      \(n\)-type for some finite \(n\).
  \end{enumerate}
\end{romdefinition}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Split Enumerable}
\begin{romdefinition}[Split Enumerable Set] \label{split-enum-def}
  \begin{equation} \label{split-enum-def-eqn}
    \mathcal{E}!(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    && \text{Def.~\ref{split-enum-def} }(\mathcal{E}!)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    && \text{Eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{Eqn.~\ref{sp-surj-eqn} (sp-surj)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{Def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    && \text{Eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    && \text{Eqn.~\ref{sp-surj-arrow-eqn} } (\twoheadrightarrow!)
  \end{align*}
  \qed
\end{proof}

@@

Split enumerability implies decidable equality on the underlying type.
To prove this, we will make use of the following lemma, proven in the
formalisation:
\begin{romlemma} \label{discrete-surj}
  \begin{equation}
    \frac{
        A \twoheadrightarrow! \; B \; \; \; \text{Discrete}(A)
      }{
       \text{Discrete}(B) 
      }
  \end{equation}
\end{romlemma}
\begin{romlemma}
  Every split enumerable type is discrete.
\end{romlemma}
\begin{proof}
  Let \(A\) be a split enumerable type.
  By lemma~\ref{split-enum-is-split-surj}, there is a surjection from
  \(\mathbf{Fin}(n)\) for some \(n\).
  Also, we know that \(\mathbf{Fin}(n)\) is discrete (proven in our
  formalisation).
  Therefore, by lemma~\ref{discrete-surj}, \(A\) is discrete.
  \qed
\end{proof}
\subsection{Manifest Bishop}
\begin{romdefinition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
  \end{equation}
\end{romdefinition}
The only difference between manifest Bishop finiteness and split enumerability
is the membership term: here we require unique membership (\(\inunique\)),
rather than simple membership (\(\in\)).

@@

% Before giving the definition of the predicate, we will first define lists (and
% membership thereof): we have chosen a \emph{container}-based definition for this
% work.

@@

% Finally, we can define formally split enumerability.
% \subsection{Split Surjections}
% \begin{romtheorem} \label{split-enum-surj}
%   Split enumerability is equivalent to a split surjection from a finite prefix
%   of the natural numbers.
%   \begin{equation}
%     \mathcal{E}!(A) \iff \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   After sufficient inlining, it emerges that our goal is simply a reassociation.
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-surj}
%   To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
%   have written out the extra lines for clarity alone.
% \end{proof}
% \begin{romlemma} \label{split-enum-discrete}
%   Any split enumerable type has decidable equality (is discrete).
% \end{romlemma}
% \begin{proof}
%   We use a corollary that if there is a split-surjection from \(A\) to \(B\),
%   and \(A\) is discrete, then \(B\) is also discrete.
% \end{proof}
% By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
% this implies split enumerable types must all be sets.
% \subsection{Closure}
% In this section we will prove closure under various operations for split
% enumerable sets.
% We are working towards a topos proof, which requires us to prove closure under
% a variety of operations: for now, we only have enough machinery to demonstrate
% the semiring operations, and dependent sums.
% In order to show closure under exponentials (function arrows), we will need an
% equivalence with \(\mathbf{Fin}\), which will be provided in
% section~\ref{manifest-bishop}.
% \begin{romlemma} \label{split-enum-bool-top-bot}
%   \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
% \end{romlemma}
% \begin{proof}
%   Each of these types clearly has a finite number of elements (2, 1, and 0,
%   respectively), and furthermore has a straightforward enumeration.
% \end{proof}
% \begin{romtheorem} \label{split-enum-sigma}
%   Split-enumerability is closed under \(\Sigma\).
%   \begin{equation} \label{split-enum-sigma-clos-sig}
%     \frac{
%       \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
%     }{
%       \mathcal{E}!(\Sigma (x : A) , U(x))
%     }
%   \end{equation}
% \end{romtheorem}
% \begin{proof}
%   To obtain the support list, we concatenate the support lists of all the proofs
%   of split-finiteness for \(U\) over the support list of \(E_A\).
%   In Agda:
%   \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-sup}
%   ``do-notation'' is available to us as we're working in the list monad.
% \end{proof}
% Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
% as these types can be defined in terms of \(\Sigma\).
% \begin{align}
%   A \times B &= \Sigma {(x : A)} , B \\
%   A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
% \end{align}
% \section{Manifest Bishop Finiteness} \label{manifest-bishop}
% Where split enumerability was the enumeration form of a split surjection from
% \(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
% \emph{equivalence} with \(\mathbf{Fin}\).
